Order statistics and numerical optimization: the case of a single observation

### 2023-10-31 inst/NBA.Rmd
consistentification

---
title: "An analysis of a single basketball game"
output: html_document
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

File `NBA.txt` contains a point-by-point analysis of a basketball
match between the Cleveland Cavaliers and the Golden State Warriors on
13 June 2017, data taken from
[here](https://www.espn.com/nba/playbyplay?gameId=400954514).

>>>>

---
title: "An analysis of a single basketball game"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I consider a single basketball match from the perspective of the
Bradley-Terry model.  File `NBA.txt` contains a point-by-point
analysis of a basketball match between the Cleveland Cavaliers and the
Golden State Warriors on 13 June 2017, data taken from
[here](https://www.espn.com/nba/playbyplay?gameId=400954514).  Some
further documentation is given at `inst/NBA.Rd` in the package.

### 2023-10-31 inst/T20.Rmd
consistentification

---
title: "The impact of winning the coin toss on the Indian Premier League"
author: "Tanvi Chandel and Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{T20}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

# The Indian Premier League

This script creates and analyses hyper2 object `T20` which is a
likelihood function for the strengths of the competitors in the Indian
Premier League.

@@

Dataframe `T20_table` has one row for each T20 IPL match in the
period 2008-2017 with the exception of three no-result matches and
seven tied matches, which were removed.  Object `T20` is a
likelihood function for the strengths of the 13 teams, and
`T20_toss` is a likelihood function that also includes a toss
strength term.

some details are given in the package help file
`T20.Rd` (type `?T20` at the R prompt).

>>>>

---
title: "The impact of winning the coin toss on the Indian Premier League"
author: "Tanvi Chandel and Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This script creates and analyses hyper2 object `T20` which is a
likelihood function for the strengths of the competitors in the Indian
Premier League.
Dataframe `T20_table` has one row for each T20 IPL match in the
period 2008-2017 with the exception of three no-result matches and
seven tied matches, which were removed.

@@

Object `T20` is a likelihood function for the strengths of the 13
teams, and `T20_toss` is a likelihood function that also includes a
toss strength term.  Some details are given in the package help file
`T20.Rd` (type `?T20` at the R prompt).

### 2023-10-31 inst/cook_repeated.Rmd
consistentification

---
title: "Judging domestic cooking ability: repeated incomplete rank observations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

(takes about five minutes without cache).
Suppose there are five people in a family.  Each day for five days,
one person cooks a meal.  After the fifth day, each person considers
the four meals cooked by other family members, and puts them in order
of preference.  This process is repeated and here I consider how to
interpret such data.

@@

### First row of the dataset

@@

## Complete dataset

@@

# Preferential voting

>>>>

---
title: "Judging domestic cooking ability: repeated incomplete rank observations"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a;
takes about five minutes without cache.  Suppose there are five people
in a family.  Each day for five days, one person cooks a meal.  After
the fifth day, each person considers the four meals cooked by other
family members, and puts them in order of preference.  This process is
repeated and here I consider how to interpret such data.

@@

@@

## Complete dataset {-}

@@

# Preferential voting {-}

### 2023-10-31 inst/counterstrike.Rmd
use cache
consistentification

---
title: "A statistical analysis of a counterstrike game: Boston 2018"
output: html_document
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

# Abstract

In this short vignette I use the `hyper2` package to analyse an
influential Counter-Strike match between two professional teams (_FaZe
Clan_ and _Cloud9_) in the major championships at Boston 2018.  I use
likelihood methods to demonstrate that one player, Skadoodle, has a
legitimate claim to be the strongest player on the field; but the data
is indistinguishable from a null of equal player strength.
Counter-strike dataset kindly supplied by Zachary Hankin.

@@

```{r, plotit}
counterstrike_maxp <- maxp(counterstrike)
dotchart(counterstrike_maxp,pch=16,main='observed data')
```

@@

```{r, randomdata}

>>>>

---
title: "A statistical analysis of a counterstrike game: Boston 2018"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
In this short vignette I use the `hyper2` package to analyse an
influential Counter-Strike match between two professional teams (_FaZe
Clan_ and _Cloud9_) in the major championships at Boston 2018.  I use
likelihood methods to demonstrate that one player, Skadoodle, has a
legitimate claim to be the strongest player on the field; but the data
is indistinguishable from a null of equal player strength.
Counter-strike dataset kindly supplied by Zachary Hankin.

@@

```{r, calcmax,cache=TRUE}
counterstrike_maxp <- maxp(counterstrike)
```

```{r, dotchartmaxp}
dotchart(counterstrike_maxp,pch=16,main='observed data')
```

@@

```{r, label=randomdata,cache=TRUE}

### 2023-10-31 inst/distance_matrix_idea.Rmd
typo
name all chunks and cache the slow ones
consistentification

---
title: "icons: some experimental results"
output: html_document
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This file, `icons2.Rmd`, presents some speculative experimental work
which is under development.  Do not confuse this file with
`icons.Rmd`, which creates the `icons` objects in the package.

@@

```{r}
M
(M <- round(M*100))
```

@@

```{r}
plot(phangorn::upgma(as.dist(M)))
```

>>>>

---
title: "icons: some experimental results"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This file presents some _speculative_ experimental work which is under
development.  Do not confuse this file with `icons.Rmd`, which creates
the `icons` objects in the package.

@@

```{r roundm}
M
(M <- round(M*100))
```

@@

```{r plotphangorn}
plot(phangorn::upgma(as.dist(M)))
```

### References {-}

### 2023-10-31 inst/eurodance.Rmd
say how long it takes to run
consistentification

---
title: "Eurovision **Dance** Contest 2008"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

The file `eurodance2008.txt`, used below, is copied from
"Eurovision Dance Contest 2008", Wikipedia, accessed March 31, 2021.
First we specify the matrix as appearing in the Wikipedia page:

@@

>>>>

---
title: "Eurovision **Dance** Contest 2008"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a;
takes about 20 minutes to process without cache.  The file
`eurodance2008.txt`, used below, is copied from "Eurovision Dance
Contest 2008", Wikipedia, accessed March 31, 2021.  First we specify
the matrix as appearing in the Wikipedia page:

@@

###R References {-}

### 2023-10-31 inst/eurovision.Rmd
consistentification

---
title: "Eurovision Song Contest: semifinal 1, 2009"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

The file `eurovision.txt`, used below, is copied from "Eurovision Song
Contest 2009," Wikipedia, accessed May 13, 2018.  It refers to
semi-final 1.  More documentation is given in `eurovision.Rd` [type
`help(euro2009)` at the `R` prompt].

@@

# References {-}

>>>>

---
title: "Eurovision Song Contest: semifinal 1, 2009"
output: bookdown::html_document2
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
The file `eurovision.txt`, used below, is copied from "Eurovision Song
Contest 2009," Wikipedia, accessed May 13, 2018.  It refers to
semi-final 1.  More documentation is given in `eurovision.Rd` [type
`help(euro2009)` at the `R` prompt].

@@

### 2023-10-31 inst/fatigue.Rmd
consistentification

---
title: "fatigue"
author: "Robin Hankin"
date: "16/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

## R Markdown

Analysis of  the 1973 interzonal competition

>>>>

---
title: "fatigue"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Analysis of the 1973 interzonal competition.

### 2023-10-31 inst/global_liveability_ranking.Rmd
consistentification

---
title: "Global liveability ranking"
output: html_document
---

@@

@@

(takes about an hour to process without cache)

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

@@

>>>>

---
title: "Global liveability ranking"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a;
takes about an hour to process without cache.

@@

@@

## References {-]

### 2023-10-31 inst/grad_weight.Rmd
consistentification

---
title: "grad_weight"
author: "Robin KS Hankin"
date: "10/04/2022"
output: html_document
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

## Differentiate with respect to weights for a `hyper3` object

We will start with a simple example.  Consider two players $p_1$ and
$p_2$ who play tennis say, $a+b$ times with $p_1$ winning $a$ times
and $p_2$ winning $b$ times.  We use the following likelihood function:

>>>>

---
title: "Differentiation with respect to weights for a `hyper3` object"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
We will start with a simple example.  Consider two players $p_1$ and
$p_2$ who play tennis say, $a+b$ times with $p_1$ winning $a$ times
and $p_2$ winning $b$ times.  We use the following likelihood function:

### 2023-10-31 inst/handover.Rmd
consistentification

---
title: "handover"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document is based on Altham and Hankin 2010, which
discussed an interesting dataset originally presented by Lin
et. al. (2009).  It uses more sophisticated idiom furnished by the
`hyper2` package, which is an improved and efficient upgrade of the
primitive `hyperdirichlet` package.  The dataset arose from 69 medical
malpractice claims and refers to the two surgeon–reviewers' answers to
the question "was there a communication breakdown in the hand-off
between physicians caring for the patient?".  The rows of Table 1
correspond to the answers that were given by reviewer 1, and the
columns to the answers given by reviewer 2.

@@

* PME Altham and RKS Hankin 2010.  "Using recently developed software on
a 2x2 table of matched pairs with incompletely classified data".
_Journal of the Royal Statistical Society_, 59:2):377-379
* L West and RKS Hankin 2008. "Exact tests for two-way contingency
tables with structural zeros".  _Journal of Statistical Software_
28(11)
* RKS Hankin 2010. "A generalization of the Dirichlet distribution". 
 _Journal of Statistical Software_ 33(11)
* RKS Hankin 2017. "Partial rank data with the hyper2 package:
  likelihood functions for generalized Bradley-Terry models".  _The R
  Journal_ 9(2)

>>>>

---
title: "handover issues in medical care"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document is based on @altham2010, which
discussed an interesting dataset originally presented by Lin
et. al. (2009).  It uses more sophisticated idiom furnished by the
`hyper2` package, which is an improved and efficient upgrade of the
primitive `hyperdirichlet` package.  The dataset arose from 69 medical
malpractice claims and refers to the two surgeon–reviewers' answers to
the question "was there a communication breakdown in the hand-off
between physicians caring for the patient?".  The rows of Table 1
correspond to the answers that were given by reviewer 1, and the
columns to the answers given by reviewer 2.

@@

### 2023-10-31 inst/hepatitis.Rmd
consistentification

---
title: "Hepatitis"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

>>>>

---
title: "Hepatitis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.

### 2023-10-31 inst/home_advantage.Rmd
consistentification

---
title: "Quantifying the home ground advantage"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{home_ground_advantage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

@@

(takes about fifteen minutes to run without cache)

Here I analyse a small synthetic dataset for home ground advantage,
then go on to discuss a dataset presented by Agresti.  But first, the
smallest nontrivial example is a 2x2 case:

>>>>

---
title: "Quantifying the home ground advantage with `hyper2` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a
(takes about fifteen minutes to run without cache).  Here I analyse a
small synthetic dataset for home ground advantage, then go on to
discuss a dataset presented by Agresti.  But first, the smallest
nontrivial example is a 2x2 case:

### 2023-10-31 inst/home_advantage3.Rmd
consistentification

---
title: "Quantifying the home advantage with `hyper3` objects"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{hyper3}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

@@

## Aside: `hyper2` analysis of the same scoreline

>>>>

---
title: "Quantifying the home advantage with `hyper3` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

@@

# Aside: `hyper2` analysis of the same scoreline

### 2023-10-31 inst/hunter.Rmd
correct many mistakes
consistentification

---
title: "Iteration techniques for Plackett-Luce likelihood functions and the `hyper2` package"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here I discuss why implementing equation 30 of Hunter (2004) for
`hyper2` objects representing Plackett-Luce likelihood functions is
difficult.  Consider the following rank observations: `724356` [that
is, 7 came first, then 2, then 4 then 3 then 5 then 6] and `1372`
[that is, 1 came first, then 3 then 7 then 2].  This would have a
Plackett-Luce likelihood function of

@@

\[
p_3\longrightarrow\frac{p_3}{(p_7+p_2+p_4+p_3)^{-1}+(p_1+p_3)^{-1}}=
\frac{[3]}{[7243]^{-1}+[13]^{-1}}
\]

@@

$p_1 \longrightarrow [1]$

$p_2 \longrightarrow\frac{[2]}{[27]^{-1}+[1237]^{-1}}$

$p_3 \longrightarrow\frac{[3]}{[2347]^{-1}+[13]^{-1}}$

$p_4 \longrightarrow\frac{[4]}{[247]^{-1}}$

$p_5 \longrightarrow\frac{[5]}{[23457]^{-1}}$

$p_6 \longrightarrow\frac{[6]}{[234567]^{-1}}$

$p_7 \longrightarrow\frac{[7]}{[137]^{-1}}$

and it does not seem at all easy to convert the `hyper2` object to
Hunter's iteration scheme.  To update $p_3$, for example, we would
start with the denominator $[1237]\cdot[23456]\cdot[234567]\cdot
[237]\cdot[27]\cdot[3456]\cdot[356]\cdot[56]$ and then extract just
the brackets containing 3 to get $[1237]\cdot[23456]\cdot[234567]\cdot
[237]\cdot[3456]\cdot[356]$ and then somehow identify two increasing chains

@@

and _then_ somehow identify from these two chains that 1 beats 3
(leading to $[13]$ in the iteration) and also that $\left\lbrace
2,4,7\right\rbrace$ beat 3 (leading to $[2347]$).

### Reference {-}

* D. R. Hunter 2004.  "MM algorithms for generalized Bradley-Terry
  models".  _The Annals of Statistics_, 32(1):384

>>>>

---
title: "Iteration techniques for Plackett-Luce likelihood functions and the `hyper2` package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I discuss why implementing equation 30 of [@hunter2004] for
`hyper2` objects representing Plackett-Luce likelihood functions is
difficult.  Consider the following rank observations: `724356` [that
is, 7 came first, then 2, then 4 then 3 then 5 then 6] and `1372`
[that is, 1 came first, then 3 then 7 then 2].  If these two
observations are independent, we would have a Plackett-Luce likelihood
function of

@@

\[
p_3\longrightarrow\frac{2}{
(p_7+p_2+p_4+p_3+p_5+p_6)^{-1}+
(    p_2+p_4+p_3+p_5+p_6)^{-1}+
(        p_4+p_3+p_5+p_6)^{-1}+
(            p_3+p_5+p_6)^{-1}+
(p_1+p_3+p_7+p_2)^{-1}+
(    p_3+p_7+p_2)^{-1}
}\]

or, more compactly,
\[p_3\longrightarrow
\frac{2}{[724356]^{-1}+[24356]^{-1}+[4356]^{-1}+[356]^{-1}+[1372]^{-1}+[372]^{-1}}.
\]

@@

$p_1\longrightarrow\frac{1}{[1237]^{-1}}$

$p_2\longrightarrow\frac{1}{[234567]^{-1}+[23456]^{-1}}$

$p_3\longrightarrow
\frac{2}{[273456]^{-1}+[23456]^{-1}+[3456]^{-1}+[356]^{-1}+[1237]^{-1}+[237]^{-1}}$

$p_4 \longrightarrow\frac{1}{[724356]^{-1}+[24356]^{-1}+[4356]^{-1}}$

$p_5\longrightarrow\frac{1}{[273456]^{-1}+[23456]^{-1}+[3456]^{-1}+[356]^{-1}+[56]^{-1}}$

$p_6 \longrightarrow\frac{0}{[6]^{-1}}$

$p_7 \longrightarrow\frac{2}{[273456]^{-1}+[27]^{-1}}$

and it does not seem at all easy to convert the `hyper2` object to
Hunter's iteration scheme.  The numerators of the fractions are just
the powers of the singleton terms in the likelihood function.  But the
denominators are harder.  To update $p_3$, for example, we would start
with the denominator $[1237]\cdot[23456]\cdot[234567]\cdot
[237]\cdot[27]\cdot[3456]\cdot[356]\cdot[56]$ and then extract just
the brackets containing 3 to get $[1237]\cdot[23456]\cdot[234567]\cdot
[237]\cdot[3456]\cdot[356]$ and then somehow identify two increasing
chains

@@

### References {-}

### 2023-10-31 inst/hyper3.Rmd
consistentification

---
title: "The weighted hyperdirichlet distribution: the `hyper3` class"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{hyper3}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

>>>>

---
title: "The weighted hyperdirichlet distribution: the `hyper3` class"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

### 2023-10-31 inst/interzonal.Rmd
consistentification

---
title: "Analysis of the 1962 world chess championship, Curacao"
output: html_document
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This file creates likelihood functions `H` and `H_coll`, which appear in
the package as `interzonal` and `collusion` respectively.

@@

# References

C. C. Moul and J. V. C. Nye 2009. "Did the Soviets collude? A
 statistical analysis of championship chess 1940-1978.  Journal of
 Economic Behaviour and Organization 70(2009) 10-21

>>>>

---
title: "Analysis of the 1962 world chess championship, Curacao"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This file creates likelihood functions `H` and `H_coll`, which appear
in the package as `interzonal` and `collusion` respectively.  It also
presents some analysis of the datasets using the `hyper2` package.

@@

# References {-}

### 2023-10-31 inst/karate.Rmd
consistentification

---
title: "karate"
output: html_document
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here I analyse a [karate](https://en.wikipedia.org/wiki/Karate)
dataset taken from
[wikipedia](https://en.wikipedia.org/wiki/2018_World_Karate_Championships).
Short story: the dataset has too many degrees of freedom to be easily
analysed with `hyper2` and Zermelo's iteration scheme as implemented
by the `BradleyTerry` package exploits the special pairwise structure
of the likelihood function very effectively.

>>>>

---
title: "karate"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I analyse a [karate](https://en.wikipedia.org/wiki/Karate)
dataset taken from
[wikipedia](https://en.wikipedia.org/wiki/2018_World_Karate_Championships).
Short story: the dataset has too many degrees of freedom to be easily
analysed with `hyper2` and Zermelo's iteration scheme as implemented
by the `BradleyTerry` package exploits the special pairwise structure
of the likelihood function very effectively.

### 2023-10-31 inst/kka.Rmd
consistentification

---
title: "Karpov, Kasparov, Anand"
output: html_document
---

@@

(takes about 45 minutes to process without cache)

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This file defines dataset `kka` which is the observed scorelines
between players Karpov, Kasparov, and Anand.  It is used to calculate
likelihood functions: `karpov_kasparov_anand`, `kka_3whites` and
`kka_3draws`.  This Rmd file supercedes files
`inst/karpov_kasparov_anand.R` and `inst/kka_3draws.R` and
`inst/kka_array.R`.  Some more detail is given in
`man/kka.Rd`.

>>>>

---
title: "A Bradley-Terry analysis of three grandmasters: Karpov, Kasparov, Anand"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a;
this `Rmd` file takes about 45 minutes to process without cache.  Here
we define dataset `kka` which is the observed scorelines between
players Karpov, Kasparov, and Anand.  It is used to calculate
likelihood functions: `karpov_kasparov_anand`, `kka_3whites` and
`kka_3draws`.  This Rmd file supercedes files
`inst/karpov_kasparov_anand.R` and `inst/kka_3draws.R` and
`inst/kka_array.R`.  Some more detail is given in `man/kka.Rd`.

### 2023-10-31 inst/monster_vs_lambda.Rmd
consistentification

---
title: "Monster or Lambda?  A likelihood analysis"
author: "Robin KS Hankin"
output: html_document
---

@@

(takes about an hour to process without cache)

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here I consider pairwise comparisons with a home-ground advantage and
wish to compare two likelihood functions.  First, the monster parametrization:

>>>>

---
title: "Monster or Lambda?  A likelihood analysis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a;
takes about an hour to process without cache.  Here I consider
pairwise comparisons with a home-ground advantage and wish to compare
two likelihood functions.  First, the monster parametrization:

### 2023-10-31 inst/multiset_hyper2.Rmd
consistentification

---
title: "The `cheering3()` probability model for order statistics: a
verification with the `multiset()` function of the `partitions` package"
output: html_document
date: "2022-07-28"
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here we will sum probabilities across a sample space and verify that
the total is constant.  In theory the sum of the probabilities of all
members of the sample space is 1, but in practice the likelihood
calculation includes an indeterminate constant which means that all we
can say is that the total probability is not a function of the
parameters of the distribution.

@@

Above we see very tight clustering about the mean value of 0.25

>>>>

---
title: "The `cheering3()` probability model for order statistics: a
verification with the `multiset()` function of the `partitions` package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Here we will sum probabilities across a sample space and verify that
the total is constant.  In theory the sum of the probabilities of all
members of the sample space is 1, but in practice [and certainly in
`hyper2` idiom], the likelihood calculation includes an indeterminate
constant which means that all we can say is that the total probability
is not a function of the parameters of the distribution.

Here I present some relatively informal and unstructured thoughts on
this issue from a numerical and `hyper2` perspective.

@@

Above we see very tight clustering about the mean value of 0.25.

### References {-}

### 2023-10-31 inst/order.Rmd
consistentification

---
title: "Order statistics and numerical optimization: the case of a single observation"
output: html_document
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

## The `hyper2` package

We consider the likelihood function given by a single observation of
an order statistic.  Here we investigate some numerical problems
arising from the use of `optim()` and `constrOptim()`.  We will
consider $n=10$ competitors and without loss of generality, the
competitors finish in order $1,2,3,\ldots,10$.  The likelihood
function for this observation is:

>>>>

---
title: "Order statistics and numerical optimization: the case of a single observation"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I consider the likelihood function given by a single observation
of an order statistic, which is problematic in the sense that the
maximum likelihood estimate for the BT strengths of the competitors is
$(1,0,\ldots,0)$.  Here we investigate some numerical problems arising
from the use of `optim()` and `constrOptim()`.  We will consider
$n=10$ competitors and without loss of generality, the competitors
finish in order $1,2,3,\ldots,10$.  The likelihood function for this
observation is:
